Pre-orders or Philanthropy: An Analysis of Altruism on Kickstarter and Other Crowdfunding Sites

### 2014-12-19 AltruisticLanguage/writing/paper.tex
editing done

\def\year{2015}
%File: formatting-instruction.tex
\documentclass[letterpaper]{article}
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{url}
\usepackage{array}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subfig}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newenvironment{blockquote}{%
  \setlength{\parskip}{.5em}
  \par%
  \small
  \medskip
  \leftskip=3em\rightskip=3em%
  \noindent\ignorespaces}{%
  \par\medskip}
\newcommand{\specialcell}[2][l]{%
  \begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
\frenchspacing
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\pdfinfo{
/Title Pre-orders or Philanthropy: An Analysis of Altruism on Kickstarter
/Author Jack Hessel
/Keywords X,Y,Z}
\setcounter{secnumdepth}{0}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Pre-orders or Philanthropy: An Analysis of Altruism on Kickstarter and Other Crowdfunding Sites}
\author{Jack Hessel}
\maketitle
\begin{abstract}
\begin{quote}
In this study we conduct several experiments related to altruistic behavior on Kickstarter and other crowdfunding sites. This realm has been relatively unexplored in previous work, which is surprising given that roughly 20\% of all backing money on Kickstarter is donated. Because previous work sugguests that lingustic features are very important to requests made online, language is a focal point of our analysis. We first find that Kickstarter exists in a lingustic realm somewhere between pure altruism and pure commericalism. Next, we find that useful information about the sequential nature of requests can be learned in an unsupervized fashion. Also, we find that linguistic features on Kickstarter are able to predict whether or not projects recieve above-average levels of prosocial activity. Finally, and most surprisingly, we find that failing projects are acutally more likely to recieve altruistic donations, sugguesting the existance of a subset of users who frequently give small amounts of money to failing projects and request no rewards in return.
\end{quote}
\end{abstract}

@@

me & 80.57 & would be & 37.02 & take care of & 16.27 &
perfect & 64.77 & is the & 48.67 & the united states & 2.88 \\\hline
was & 82.01 & name is & 37.74 & to help with & 17.72 &
has & 73.00 & and is & 54.69 & is a great & 3.75\\\hline
am & 82.62 & family and & 40.60 & in need of & 19.04 &
with & 78.24 & with a & 56.31 & is one of & 27.54\\\hline
family & 98.81 & thank you & 41.80 & thank you for & 20.38 &
includes & 81.00 & has a & 56.71 & of the most & 30.88\\\hline
i & 136.42 & to help & 55.71 & friends and family & 27.25 &
easy & 82.72 & out of & 71.67 & one of the & 30.97 \\\hline
\end{tabular}
\caption{Summary of Bayesian analysis of language comparing Kickstarter to GoFundMe and Amazon. Negative Z-scores correspond to Kickstarter. Top ten most significant n-grams are presented.}
\label{tab:pairwise}
\end{table*}

Furthermore, in our dataset consisting of 45K completed Kickstarter projects, we are able to quantify the number of purely altruistic donation events and the total value given above and beyond the amount required for all reward claims. Of the roughly 3.4M donation events, about 360K (10.59\%) resulted from users who selected no reward. Of the roughly \$246M pledged in our dataset, over \$53M (21.59\%) was given altruistically.

The community and language of Kickstarter have been studied previously. Mitra and Gilbert analyze the language that predicts successful Kickstarter projects \cite{mitra2014language}. When compared their baseline regression trained on language-indepedent project features, their final classifier, which includes n-gram counts, is able to classify projects 15\% more accurately, resulting in a 97.8\% correct prediction rate in a 10-fold cross validation setting. Similar analyses that do not take language fully into account are not able to achieve such predictive accuracy. Studies that use of minimal textual information in conjunction with baseline project features are able to learn classifiers that reach 68\% accuracy using information available at launch time \cite{greenberg2013crowdfunding}. Even using post-launch information, however, does not improve accuracy as much as accounting for language; 85\% accuracy is achieved using information available after 15\% of the duration of the campaign has expired \cite{etter2013launch}. Furthermore, lingustic factors have been shown to effect the quality and time-to-response of questions posed in an online community \cite{teevan2011factors}. These findings sugguest that language is very important on Kickstarter, which is a primary motivation for making it a focus of our analysis.

Altruism in online communities has also been studied as well. Althoff et al. examine a purely altruistic community wherein users request free pizza from other members \cite{althoff2014ask}. An advantage of their dataset is that all altruistic requests are made for the same service. Previous work on Kickstarter has also demonstrated that users are more likely to donate to campigns as time is winding down and the project has not yet met its goal, which is an arguably altruistic behavior \cite{kuppuswamy2013crowdfunding}.

@@

\section{Primary Dataset}
Our dataset consists of 45810 completed Kickstarter projects, the URLs of which were originally collected by AppsBlogger.com. The projects in our data were all active sometime between April 2009, the earliest start date, and August 2012, the latest end date. Every Kickstarter project has an associated HTML page, which we parse using the popular HTML parsing library Beautiful Soup 4\footnote{\url{http://www.crummy.com/software/BeautifulSoup/}}. Each project's page contains information about the project such as its title, its description, the rewards (with associated cost, number of backers who have claimed that reward, and a description), the FAQ section, the total number of backers, and the total amount of money raised. This information served as the basis of our analysis.

Kickstarter projects are also associated with a single category. The category of a Kickstarter project is used by the site to index projects and to help users find projects of a specific type. While some of these categories (13) are ``main'' categories, most are ``subcategories'' that have a corresponding main category. In most cases, we treat all categories as equivalent, irrespective of whether or not it is a sub or a main.

In all of our analyses as a preprocessing step for text data, we remove punctuation from strings and convert all characters to lower case, but do not discard stop words. In general, we are interested in unigrams, bigrams, and trigrams, and tokenize after these preprocessing steps.

@@

Amazon is a very popular online marketplace, where users can purchase items from a wide range of categories. Each product is associated with a description and and a set of categories. Our dataset consists of just over 1M Amazon product descriptions \cite{mcauley2013hidden}. We believe that Amazon product descriptions can be considered as purely commercial. On the other hand, GoFundMe is a crowdfunding site similar to Kickstarter, but campaigns are generally more focused on charitable/personal ventures. For instance, a popular category on GoFundMe is ``Medical,'' wherein users create pages to raise money for medical expenses for family and friends. Our dataset consists of 7.7K GoFundMe project descriptions scraped in November 2014. In contrast to Kickstarter, in a vast majority of cases GoFundMe contributors don't have the option to select compensentory rewards, and we therefore regard our GoFundMe data as requests for purely altruistic actions.

To compare the language used on Kickstarter to purely commercial and purely altrustic pitches, we adopt a Bayesian approach that has previously been successful in comparing the language used by two political parties on a specific issue \cite{monroe2008fightin}. This model has been used to compare word usage between two corpora in a princpled way. Under the model, language is modeled as a ``bag of words'' represented by a multinomial distribution over vocabulary items. For the two corpora in question, we assume that their language vectors are drawn from a Dirichlet prior. Though it is possible in this scenario to encode useful information in the prior, we find it sufficient to utilize an uninformative prior and set our Dirichlet hyperparameter to be $\alpha=.01$. Assuming this language model, it becomes possible to compute approximate differences in vocabulary usages, and approximate variances of those differences.

In order to use this model, one must first define a vocabulary. Because we are interested in language rather than specific subject matter, we apply a cross-categorical filtering scheme in an attempt to control for topic. Recall that in each of our datasets, documents are associated with one or more categories. In a first attempt to control for topic, we first apply a filter that removes all n-grams that don't appear in all categories at least once. We found this simple filtration scheme to be inadequate -- the words ``book,'' ``publisher,'' and ``author,'' for instance, appear very frequently in the Amazon dataset because about 78\% of the products in the Amazon dataset are tagged under the Book category.

To apply a more stringent topic controlling filter after removing n-grams that don't appear in all categories, for each n-gram $n$, we compute the intra-category document frequency denoted $f_n^c$ for all categories $c$. We then compute the ``normalized width'' $w_n$ for n-gram $n$ as...

@@

The intuition behind normalized width is as follows -- if a word appears in all categories with roughly the same frequency, the normalized width will be small. If a word appears very frequently in a single category, the normalized width will be large. After computing all normalized widths, we filter out the 10\% of n-grams with highest normalized width.

After this two-part filtration scheme is applied to all three datasets, we take an intersection of the resulting n-grams to form our final vocabulary, which contains 2077 n-grams. The results of these analyses are summarized in Table \ref{tab:pairwise}.

@@

To emperically analyze these n-gram comparisons, we first rank the n-grams in each experiment by their z-scores and compute Kendall's tau correlation between the Kickstarter vs. Amazon and Kickstarter vs. GoFundMe analyses. The intuition behind this comparsion is as follows: if Kickstarter is more altruistic than Amazon, but less altruistic than GoFundMe, then we would expect altruistically-associated n-grams to be on opposite ends of each ranking. This phenomenon occurs

@@

To analyze the sequential structure of requests in a crowdfunding setting, we run a case study on a subset of data utilizing a content model \cite{barzilay2004catching} to produce a sentence clustering. A content model is an unsupervized learning algorithm that operates on the sentence level, takes information order into account, and learns transition probabilities between sentence types. Here we provide a brief overview of the content model algorithm; for a fuller description, consult the original paper.

The goal of a content model is to learn a set of sentence clusters utilizing ordering information. Given a set of $N$ documents where document $n$ has $s_n$ sentences, sentences are first clustered into $k$ groups according to bigram features using complete-link clustering with a cosine distance metric. After sentences are clustered into these groups, we merge all clusters smaller than some threshold into an ``etcetera'' cluster. This etcetera cluster is a unique feature of the content model that allows it the flexibility to sort sentences that otherwise don't fit into an ``other'' cluster. After this merging process, we are left with $m$ clusters.

@@

After training these language models, we now treat each clustering as a state in a Hidden Markov Model, and all input documents as emissions from that Hidden Markov Model. We are able to compute the Viterbi path for each document, and we re-cluster sentences based on these most-likely paths. Finally, we retrain each of the language models based on the new clustering. This process of re-clustering, computing Viterbi paths, and re-training language models in an expectation 
maximization like fashion is executed until convergence or a maximum number of iterations is made.

@@

Clearly, there are clusters of all sorts that represent different pieces of a plea for help with medical expenses. One interesting cluster is the ``Ability'' cluster, which contains 356 sentences. Describing in detail a lack of ability is likely to elicit empathetic feelings from a reader, and, according to the well-supported empathy-altruism hypothesis, empathetic feelings are very likely to lead one to purely altruistic behavior, such as donating on a crowdfunding site \cite{batson1988five}.

Content models carry with them the advantage that sentences with disjoint bigram features may be clustered together based on the order of sentences. For instance, ``your contributions big or small are greatly appreciated,'' ``as a family we thank you,'' and ``god bless you have a wonderful day'' are clustered together in the ``Thank you'' cluster. Despite the three sentences having disjoint bigram features and being entirely dissimilar according to the basic intialization scheme, the content model is able to discern that the three likely have similar meaning based on their placement in documents relative to other sentences.

For the remaineder of this analysis, we are going to focus on the cluster containing 254 sentences (dubbed the ``Thank you'' cluster) and the cluster containing 272 sentences (dubbed the ``Introduction'' cluster). Though there are many interesting things that can be learned from the transition probabilities output by the content model, the most interpretable fact is illustrated in Figure \ref{fig:docpos}. This figure shows the distribution of normalized sentence position of each cluster in the sentences in the dataset. For a given cluster $c$, this statistic is defined as

\begin{equation} \label{eq:sentpos}
\sum\limits_{k \in \text{Occurances of $c$}} \frac{\text{Position of $k$ in containing sentence}}{\text{Length of sentence containing $k$}}
\end{equation}

While this result is somewhat unsurprising (situations are introduced before authors thank readers) it demonstrates that the content model is able to learn useful information about ordering in an unsupervised fashion, and that aspects of the sequential nature of altruistic requests are learnable by this method.

@@

How does altruism manifest on Kickstarter? When a user backs a project on Kickstarter (which is impossible to do anonymously) they are prompted to select a backer reward of value less than or equal to their donation amount. If a user does not want to claim a reward but, rather, donate altruistically, there exists an option that allows them to claim no rewards. Because the total number of claims for each non-altruistic reward are displayed publicly on the project page, along with the total number of backers (altruistic and non-altruistic alike) we are able to compute the total number of purely altruistic donation events for each project. Furthermore, the site allows for users to give an amount exceeding the cost of a given reward, so we are able to compute the total amount donated to a project above and beyond the amount required for all the posted reward claims. As previously noted, altruistic donation events constitute 10.59\% of all donation events and 21.59\% of all money donated in our dataset (over 53M dollars) was given away.

@@

We use glmnet \cite{friedman2010glmnet} for our regression tasks. glmnet is a software package that implements cyclic coordinate decent and is capable of performing penalized regression tasks on highly sparse datasets. For our experiments, we utilize an $L_1$ (lasso) regularization term in an attempt to enforce sparsity and derive a parsimonious model, and set glmnet to perform 10-fold cross validation to search for the optimal regularization coefficient and prevent overfitting. In addition to training a model on all the data, as a seperate measure we also report test accuracy when spliting our data into a 40k project training set and a 5k project test set. Notably, glmnet purposefully does not report standard errors on resulting feature coefficients because the regularization process intorudces substantial bias into the regression process.

Using a model consisting of controls only, we achieve a cross-validation accuracy of 64.53\% (6.19\% improvement over constant prediction) and a testing set accuracy of 63.80\% (5.46\% improvement). The most important positive and negative predictive features were all Kickstarter category control variables. Dance, theater, public art, performance art, and classical music ($\vec{\beta} = \langle 1.84, 1.56, 1.24, 1.04, .930 \rangle$, respectively) were the categorical variables which were the strongest positive predictors, whereas board/card games, games, comics, video games, and product design ($\vec{\beta} = -\langle 1.64, 1.23, 1.02, .862, .795 \rangle$, respectively) were the strongest negative predictors. This result indicates that product category is the most important factor in determining if a project will recieve altruistic donations, and perhaps that projects in artistic categories are more likely to recieve such Kickstarter backing. This result is fairly unsurprising -- artistic endevours are generally regarded as intrinsically valuable to society, as evidenced by the existance of organizations like the National Endowment of the Arts which has, to date, awarded more than 5B dollars for artistic endevours\footnote{\url{http://arts.gov/about-nea}}. The negatively predictive categories tend to have concretely associated products with their projects (a video game, for instance) and backers likely pre-order instead of donate, in general.

@@

Does accounting for linguistic content of a project increase our predictive accuracy? Our textual features consist of binary indicator variables for unigrams, bigrams, and trigrams in the project descriptions, FAQ question/answers, and reward descriptions of our Kickstarter projects. We apply the same two-part filter that we use in our previous analysis to determine a set of topic-free n-grams, which forces selected n-grams to appear in every Kickstarter category and be relatively uniformly distributed between them (see Equation \ref{eq:width}). Applying this two-part filtration process on our Kickstarter dataset exclusively, we end up with 31281 filtered n-grams.

When we incorporate these particular lingustic features, our cross-validation accuracy increases to 67.02\% (2.49\% improvement over controls-only baseline) and our testing accuracy increases to 66.68\% (2.88\% improvement). The inclusion of lingustic features, therefore, causes our predictive accuracy to increase slightly but significantly.

@@

To address this question, we first run a new lasso penalized regression task where predictive variables consist of the control variables highlighted in Table \ref{tab:controls} plus two additional binary variables ``success,'' which indicates whether or not the funding goal of the project was reached, and ``numBackers,'' the total number of backers for a given project. The target of this new regression is the proportion of altruistic donations a project recieves. Because our target is a ratio that is highly sensitive to noise if the number of project backers is small, we restrict our consideration to the 30858 projects in our dataset with 10 or more backers (though the results presented here are robust to reasonable changes in this threshold).

Surprisingly, our resulting regression assigns negative weights to both our new predictive variables. With respect to the number of backers, the negative weight assigned to this variable indicates that more popular projects are less likely to recieve altruistic donations, in general. More surprising, however, is that after controling for all variables in Table \ref{tab:controls} and project size, successful projects recieve \emph{less} altruistic donations than unsuccessful projects.

@@

When we change the focus of our analysis from events to dollar values, however, this pattern becomes less apparent, or reverses altogether. In the ``Film and Video'' category, for instance, we observe a total reversal, a significant result ($p << .001$) that states successful projects receive more dollars altruistically than failed projects. In the ``Music'' category, the test statistic is close to zero,  In fact, over all categories all statistically significant results ( state that successful projects. Sample plots of the dollar value distributions are presented in the second row of Figure \ref{fig:failsucc}.

>>>>

\def\year{2015}
%File: formatting-instruction.tex
\documentclass[letterpaper]{article}
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{url}
\usepackage{array}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subfig}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newenvironment{blockquote}{%
  \setlength{\parskip}{.5em}
  \par%
  \small
  \medskip
  \leftskip=3em\rightskip=3em%
  \noindent\ignorespaces}{%
  \par\medskip}
\newcommand{\specialcell}[2][l]{%
  \begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
\frenchspacing
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\pdfinfo{
/Title Pre-orders or Philanthropy: An Analysis of Altruism on Kickstarter
/Author Jack Hessel
/Keywords X,Y,Z}
\setcounter{secnumdepth}{0}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Pre-orders or Philanthropy: An Analysis of Altruism on Kickstarter and Other Crowdfunding Sites}
\author{Jack Hessel}
\maketitle
\begin{abstract}
\begin{quote}
In this study we conduct several experiments related to altruistic behavior on Kickstarter and other crowdfunding sites. This realm has been relatively unexplored in previous work, which is surprising given that roughly 20\% of all backing money on Kickstarter is donated. Because previous studies sugguest that lingustic features are very important to requests made online, language is a focal point of our analysis. We first find that Kickstarter exists in a lingustic realm somewhere between pure altruism and pure commericalism. Next, we find that useful information about the sequential nature of requests can be learned in an unsupervized fashion. Also, we find that linguistic features of Kickstarter campaigns are able to predict whether or not projects attract above-average levels of prosocial activity. Finally, and most surprisingly, we find that failing projects are acutally more likely to recieve altruistic donations, sugguesting the existance of a subset of users who frequently give small amounts of money to failing projects and request no rewards in return.
\end{quote}
\end{abstract}

@@

me & 80.57 & would be & 37.02 & take care of & 16.27 &
perfect & 64.77 & is the & 48.67 & the united states & 2.88 \\\hline
was & 82.01 & name is & 37.74 & to help with & 17.72 &
has & 73.00 & and is & 54.69 & is a great & 3.75\\\hline
am & 82.62 & family and & 40.60 & in need of & 19.04 &
with & 78.24 & with a & 56.31 & is one of & 27.54\\\hline
family & 98.81 & thank you & 41.80 & thank you for & 20.38 &
includes & 81.00 & has a & 56.71 & of the most & 30.88\\\hline
i & 136.42 & to help & 55.71 & friends and family & 27.25 &
easy & 82.72 & out of & 71.67 & one of the & 30.97 \\\hline
\end{tabular}
\caption{Summary of Bayesian analysis of language comparing Kickstarter to GoFundMe and Amazon. Negative Z-scores correspond to Kickstarter. Top ten most significant n-grams are presented. A full list of n-grams and z-scores is provided.\footnote{PUT LINK HERE}}
\label{tab:pairwise}
\end{table*}

Furthermore, in our dataset consisting of 45K completed Kickstarter projects, we are able to quantify the number of purely altruistic donation events and the total value given above and beyond the amount required for all reward claims. Of the roughly 3.4M donation events, about 360K (10.59\%) resulted from users who willingly opted to recieve no reward. Of the roughly \$246M pledged in our dataset, over \$53M (21.59\%) was given altruistically.

The community and language of Kickstarter have been studied previously. Mitra and Gilbert analyze the language that predicts successful Kickstarter projects \cite{mitra2014language}. When compared their baseline regression trained on a set of language-indepedent set of project features, their final classifier, which includes binary n-gram indicator variables, is able to classify projects 15\% more accurately, resulting in a 97.8\% correct prediction rate in a 10-fold cross validation setting. Similar analyses that do not take language fully into account are not able to achieve such predictive accuracy. Studies that use of minimal textual information in conjunction with baseline project features are able to learn classifiers that reach 68\% accuracy using information available at launch time \cite{greenberg2013crowdfunding}. Even using post-launch information, however, does not improve accuracy as much as accounting for language; 85\% accuracy is achieved using information available after 15\% of the duration of the campaign has expired \cite{etter2013launch}. Furthermore, lingustic factors have been shown to effect the quality and time-to-response of questions posed in an online community \cite{teevan2011factors}. These findings sugguest that language is very important on Kickstarter and, more generally, in online communities as a whole, which is a primary motivation for making it a focus of our analysis.

Altruism in online communities has also been previously studied. Althoff et al. examine a purely altruistic community wherein users request and provide free pizza to and from other community members \cite{althoff2014ask}. An advantage of their dataset is that all their request are purely altruistic, and are made for the same service, though the total donation value of their dataset is lower than that presented here. Previous work on Kickstarter has also demonstrated that users are more likely to donate to campigns as time is winding down and the project has not yet met its goal, which is an arguably altruistic behavior \cite{kuppuswamy2013crowdfunding}.

@@

\section{Primary Dataset}
Our dataset consists of 45810 completed Kickstarter projects, the URLs of which were originally collected by AppsBlogger.com. The projects in our data were all active between April 2009, the earliest start date, and August 2012, the latest end date. Every Kickstarter project has an associated HTML page, which we parse using the popular HTML parsing library Beautiful Soup 4\footnote{\url{http://www.crummy.com/software/BeautifulSoup/}}. Each project's page contains information about the project such as its title, its description, the rewards (with associated cost, number of backers who have claimed that reward, and a description), the FAQ section, the total number of backers, and the total amount of money raised. This information served as the basis of our analysis.

Kickstarter projects are also associated with a single category. The category of a Kickstarter project is used by the site to index campaigns and to help users find projects of a specific type. While some of these categories are ``main'' categories (13/49) most are ``subcategories'' that have a corresponding main category. In most cases, we treat all categories as equivalent, irrespective of whether or not it is a sub or a main.

In all of our analyses that use text data, we preprocess by removing punctuation from strings and converting all characters to lower case, though we do not discard stop words. In general, we are interested in unigrams, bigrams, and trigrams, and tokenize after these preprocessing steps.

@@

Amazon is a very popular online marketplace, where users can purchase items from a wide range of categories. Each product is associated with a description and and a set of categories. Our dataset consists of just over 1M Amazon product descriptions \cite{mcauley2013hidden}. We consider Amazon product descriptions as a gold standard purely commercial dataset.

On the other hand, GoFundMe is a crowdfunding site similar to Kickstarter, but campaigns are generally more focused on charitable/personal ventures. For instance, a popular category on GoFundMe is ``Medical,'' wherein users create pages to raise money for medical expenses for family and friends. Our dataset consists of 7.7K GoFundMe project descriptions scraped in November 2014. In contrast to Kickstarter, in a vast majority of cases GoFundMe contributors don't have the option to select compensentory rewards, and we therefore regard our GoFundMe data as requests for purely altruistic actions.

To compare the language used on Kickstarter to purely commercial and purely altrustic pitches, we adopt a Bayesian approach that has previously been successful in comparing the language used by two political parties on a specific issue \cite{monroe2008fightin}. In a general sense, this model has been used to compare word usage between two related corpora in a princpled way. Under the model, language is modeled as a ``bag of words'' represented by a multinomial distribution over vocabulary items. For the two corpora in question, we assume that their language distributions are drawn from a Dirichlet prior. Though it is possible in this scenario to encode useful information in the prior, we find it sufficient to utilize an uninformative prior and set our Dirichlet hyperparameter to be $\alpha=.01$. Assuming this language model, it becomes possible to compute approximate differences in vocabulary usages, and approximate variances of those differences.

In order to use this model, one must first define a vocabulary. Because we are interested in general language rather than specific subject matter, we apply a cross-categorical filtering scheme in an attempt to control for topic. Recall that in each of our datasets, documents are associated with one or more categories. In a first attempt to control for topic, we first applied a filter that removed all n-grams that didn't appear in all categories at least once. We found this simple filtration scheme to be inadequate -- the words ``book,'' ``publisher,'' and ``author,'' for instance, appear very frequently in the Amazon dataset because about 78\% of the products in the Amazon dataset are tagged under the Book category.

To apply a more stringent topic controlling filter, after removing n-grams that don't appear in all categories, for each n-gram $n$, we compute the intra-category document frequency denoted $f_n^c$ for all categories $c$ where $f^n_c$ is the average number of times $n$ appears in documents within category $c$. We then compute the ``normalized width'' $w_n$ for n-gram $n$ as...

@@

The intuition behind normalized width is as follows -- if a word appears in all categories with roughly the same frequency, the normalized width will be small. If a word appears very frequently in a single category, the normalized width will be large. After computing this statistic for all n-grams, we discard the 10\% of n-grams with highest normalized width. This two-part filtration scheme is applied to all three datasets, and to derive our final vocabulary we take a 3-way intersection of the resulting topic-controlled features from Kickstarter, Amazon, and GoFundMe. The final vocabulary contains 2077 n-grams. The results of these analyses are summarized in Table \ref{tab:pairwise}.

@@

To emperically analyze these n-gram comparisons, we first rank the n-grams in each experiment by their z-scores and compute Kendall's tau correlation between the Kickstarter vs. Amazon and Kickstarter vs. GoFundMe analyses. The intuition behind this comparsion is as follows: if Kickstarter is more altruistic than Amazon, but less altruistic than GoFundMe, then we would expect altruistically-associated n-grams to be on opposite ends of each ranking. This phenomenon occurs for the z-scores of n-grams like ``friends and family'' (-48.86 to 27.24) ``thank you for'' (-112.63 to 20.38) and ``love'' (-75.63 to 23.83) as one might expect, but does it happen in general?

@@

To analyze the sequential structure of altruistic requests in a crowdfunding setting, we run a case study on a subset of data utilizing a content model \cite{barzilay2004catching} to produce a sentence clustering. A content model is an unsupervized learning algorithm that operates on the sentence level, takes information order into account, and learns transition probabilities between sentence types. Here we provide a brief overview of the content model algorithm; for a fuller description, consult the original paper.

The goal of a content model is to learn a set of sentence clusters utilizing ordering information. Given a set of $N$ documents where document $n$ has $s_n$ sentences, sentences are first clustered into $k$ groups according to bigram features using complete-link clustering with a cosine distance metric. After sentences are clustered into these groups, we merge all clusters smaller than some threshold into an ``etcetera'' cluster. This etcetera cluster is a unique feature of the content model that allows it the flexibility to assign sentences that otherwise don't fit into an ``other'' cluster. After this merging process, we are left with $m$ clusters.

@@

After training these language models, we now treat each clustering as a state in a Hidden Markov Model, and all input documents as emissions from that Hidden Markov Model. We are able to compute the Viterbi path for each document, and we re-cluster sentences based on these most-likely paths. Finally, we retrain each of the language models based on the new clustering. This process of re-clustering, computing Viterbi paths, and re-training language models in an expectation 
maximization-like fashion is executed until convergence or a maximum number of iterations is made.

@@

Clearly, there are clusters of all sorts that represent different pieces of a generalized plea for help with medical expenses. One interesting cluster is the ``Ability'' cluster, which contains 356 sentences. Describing in detail a lack of ability is likely to elicit empathetic feelings from a reader, and, according to the well-supported empathy-altruism hypothesis, empathetic feelings are very likely to lead one to purely altruistic behavior, such as donating on a crowdfunding site \cite{batson1988five}.

Content models carry with them the advantage that sentences with disjoint bigram features may be clustered together based on the order of sentences after EM is applied. For instance, ``your contributions big or small are greatly appreciated,'' ``as a family we thank you,'' and ``god bless you have a wonderful day'' are clustered together. Despite the three sentences having disjoint bigram features and being entirely dissimilar according to the original, order-unaware clustering scheme, the content model is able to discern that the three likely have similar meaning based on their placement in documents relative to other sentences.

For the remaineder of this analysis, we are going to focus on the cluster containing 254 sentences (dubbed the ``Thank you'' cluster) and the cluster containing 272 sentences (dubbed the ``Introduction'' cluster). Though there are many interesting things that can be learned from the transition probabilities output by the content model, the most interpretable fact is illustrated in Figure \ref{fig:docpos}. This figure shows the distribution of normalized sentence position of each cluster in the sentences in the dataset. For a given cluster $c$, this statistic is defined as
\begin{equation} \label{eq:sentpos}
\sum\limits_{k \in \text{Occurances of $c$}} \frac{\text{Position of $k$ in containing sentence}}{\text{Length of sentence containing $k$}}
\end{equation}
While this result is somewhat unsurprising (situations are introduced before authors thank readers) it demonstrates that the content model is able to learn useful information about ordering in an unsupervised fashion, and that aspects of the sequential nature of altruistic requests are learnable by this method.

@@

How does altruism manifest on Kickstarter? When a user backs a project on Kickstarter (which is impossible to do anonymously) they are prompted to select a backer reward of value less than or equal to their donation amount. If a user does not want to claim a reward but, rather, donate altruistically, there exists an option that allows them to claim no reward. Because the total number of claims for each non-altruistic reward are displayed publicly on the project page along with the total number of backers (altruistic and non-altruistic alike) we are able to compute the total number of purely altruistic donation events for each project. Furthermore, the site allows for users to give an amount exceeding the cost of a given reward, so we are able to compute the total amount donated to a project above and beyond the amount required for all the posted reward claims. As previously noted, altruistic donation events constitute 10.59\% of all donation events and 21.59\% of all money donated in our dataset (over 53M dollars) was given away.

@@

We use glmnet \cite{friedman2010glmnet} for our regression tasks. glmnet is a software package that implements cyclic coordinate decent and is capable of performing penalized regression tasks on highly sparse datasets. For our experiments, we utilize an $L_1$ (lasso) regularization term in an attempt to enforce sparsity and derive a parsimonious model. Furthermore, we set glmnet to perform 10-fold cross validation to search for the optimal regularization coefficient and prevent overfitting. In addition to training a model on all the data using cross-validation, as a seperate measure of model generalizability we also report test accuracy when spliting our data into a 40k project training set and a 5k project test set. Notably, glmnet purposefully does not report standard errors on resulting feature coefficients because the regularization process intorudces substantial bias into the regression process.

Using a model consisting of controls only, we achieve a cross-validation accuracy of 64.53\% (6.19\% improvement over constant prediction) and a testing set accuracy of 63.80\% (5.46\% improvement). The most important positive and negative predictive features were all Kickstarter categorical control variables. Dance, theater, public art, performance art, and classical music ($\vec{\beta} = \langle 1.84, 1.56, 1.24, 1.04, .930 \rangle$, respectively) were the categorical variables which were the strongest positive predictors, whereas board/card games, games, comics, video games, and product design ($\vec{\beta} = -\langle 1.64, 1.23, 1.02, .862, .795 \rangle$, respectively) were the strongest negative predictors. This result indicates that product category is the most important factor in determining if a project will recieve altruistic donations, and perhaps that projects in artistic categories are more likely to recieve such Kickstarter backing. This result is fairly unsurprising -- artistic endevours are generally regarded as intrinsically valuable to society, as evidenced by the existance of organizations like the National Endowment of the Arts which has, to date, awarded more than 5B dollars for artistic endevours\footnote{\url{http://arts.gov/about-nea}}. The negatively predictive categories tend to have concretely associated products with their projects (a video game, for instance) and backers likely pre-order instead of donate, in general.

@@

Does accounting for linguistic content of a project increase our predictive accuracy? Our textual features consist of binary indicator variables for unigrams, bigrams, and trigrams in the project descriptions, FAQ question/answers, and reward descriptions of our Kickstarter projects. We apply the same two-part normalized width filter that we use in our previous analysis to determine a set of topic-free n-grams, which forces selected n-grams to appear in every Kickstarter category and be relatively uniformly distributed between them (see Equation \ref{eq:width}). Applying this two-part filtration process on our Kickstarter dataset exclusively, we end up with 31281 filtered n-grams.

When we incorporate these particular lingustic features as binary indicator variables, our cross-validation accuracy increases to 67.02\% (2.49\% improvement over controls-only baseline) and our testing accuracy increases to 66.68\% (2.88\% improvement). The inclusion of lingustic features, therefore, causes our predictive accuracy to increase slightly but significantly.

@@

To address this question, we first run a new lasso penalized regression task where predictive variables consist of the control variables highlighted in Table \ref{tab:controls} plus two additional binary variables, namely ``success,'' which indicates whether or not the funding goal of the project was reached, and ``numBackers,'' the total number of backers for a given project. The target of this new regression is the proportion of altruistic donations a project recieves. Because our target is a ratio that is highly sensitive to noise if the number of project backers is small, we restrict our consideration to the 30858 projects in our dataset with 10 or more backers (though the results presented here are robust to reasonable changes in this threshold).

Surprisingly, our resulting regression assigns negative weights to both our new predictive variables. With respect to the number of backers, the negative weight indicates that more popular projects are less likely to recieve altruistic donations, in general. More surprising, however, is that after controling for all variables in Table \ref{tab:controls} and project size, successful projects recieve \emph{less} altruistic donations than unsuccessful projects.

@@

When we change the focus of our analysis from events to dollar values, however, this pattern becomes less apparent, or reverses altogether. In the ``Film and Video'' category, for instance, we observe a total reversal, a significant result ($p << .001$) that states successful projects receive more dollars altruistically than failed projects. In fact, over all categories all statistically significant results state that successful projects recieve more funding. In the ``Music'' category, however, the test statistic is close to zero, and the difference is no longer significant as it was in the event-based analysis. Sample plots of the dollar value distributions are presented in the second row of Figure \ref{fig:failsucc}.
